\chapter{Introduction}
\label{chap:introduction}
Our project aims to develop a dynamic and interactive habit tracking application that allows users to set and monitor personal and professional goals efficiently. By leveraging automation and data insights, the app will provide a seamless experience that minimises manual input, making habit tracking smarter and more intuitive. Users will also be able to gain valuable insights into their progress through clear visualisations (such as graphs and reports), helping them stay motivated and accountable.

The application is designed as a cross-platform system, primarily for iOS users while also being accessible via the web. The app has been developed using React Native with Expo Go, utilising TypeScript and JavaScript, and backend by a MySQL database. By combining user activity tracking and geo-information, the app is able to intelligently record activities.

\defaultInstructions

\begin{expectations}
A good introduction gives the reader a good idea of what the project is about in the shortest amount of text.  A reader should be able to get through this chapter very quickly and decide whether this project is relevant to their interests.
\begin{itemize}
\item A summary of the overriding business objective of the application.  Explain what the client, the end-users of your application, or the organisation deploying it would gain from the application.  Explain this value from the perspective of the client's or users' world, not from your perspective as a developer.   Do not list features, requirements, or specifications here.  The software is a means to an end, not the objective in itself.
\item A statement of the type of system you built, including the platforms on which it can be used, and the technology stack used to build it.
\end{itemize}
This chapter does not require any sections of subsections.  In many projects, two short paragraphs will be enough.  However, you can use more if the objective cannot be explained in a single paragraph.  Note, though, that you will be developing the objectives further in the next chapter.
\end{expectations}

\begin{length}
A single page should be sufficient for this chapter.  More than two pages is too long.
\end{length}
